---
title: "Data Visualisation Lab-5"
author: "Bibek Sapkota"
output:
  html_notebook: default
  pdf_document: default
header-includes:
  - \usepackage{graphicx}
---

# Part-I 

## Basic Plot types

Task 1:Importing the library
```{r}
library(ggplot2)
```


Task 1: Making a basic dataframe 
```{r}
df <- data.frame(
  x=c(3,1,5),
  y=c(2,4,6),
  label=c("a","b","c")
)
```

Task 2:Visualizing the df dataset in point,bar,line,area,path, text ,tile and polygon graph
```{r}
p<- ggplot(df,aes(x,y,label=label))+
  xlab(NULL)+ylab(NULL)
  p + geom_point() + ggtitle("geom_point")
  p + geom_bar(stat="identity") +
  ggtitle("geom_bar(stat=\"identity\")")
  p + geom_line() + ggtitle("geom_line")
  p + geom_area() + ggtitle("geom_area")
  p + geom_path() + ggtitle("geom_path")
  p + geom_text() + ggtitle ("geom_text")
  p + geom_tile() + ggtitle( "geom_tile")
  p + geom_polygon() + ggtitle ("geom_polygon")
```

## Displaying Distribution

Task :visualizing distribution of diamond dataset depth in histogram and experimenting with the binwidth to find a revealing view.
```{r}
data("diamonds")

ggplot(diamonds, aes(x=depth)) + 
  geom_histogram() + 
  ggtitle("Default binwidth")

ggplot(diamonds, aes(x=depth)) + 
  geom_histogram(binwidth=0.1) + 
  ggtitle("Binwidth = 0.1")

ggplot(diamonds, aes(x=depth)) + 
  geom_histogram(binwidth=0.5) + 
  ggtitle("Binwidth = 0.5")

ggplot(diamonds, aes(x=depth)) + 
  geom_histogram(binwidth=1) + 
  ggtitle("Binwidth = 1")

```

## Dealing with overplotting

Task: Creating a dataset with 2000rows and columnsand then ploting it using geom_point with different shape to see overplotting of the data.
```{r}
df <- data.frame(x = rnorm(2000), y = rnorm(2000)) 
norm <- ggplot(df, aes(x, y))
norm + geom_point()           # Default pixexl sized
norm + geom_point(shape = 1)  # 1 pixel sized
norm + geom_point(shape = ".") # Pixel sized
```

Task:Using alpha blending(transparency) to make the points transparent to see the data more easily for large dataset with overplotting.
```{r}
norm + geom_point(colour = alpha("black", 1/3))
norm + geom_point(colour = alpha("black", 1/5))
norm + geom_point(colour = alpha("black", 1/10))
norm + geom_point(colour = alpha("black", 1/256)) # you can go till 1/256 transparency which result in clear graph
```


```{r}
td <- ggplot(diamonds, aes(table, depth)) +
xlim(50, 70) + ylim(50, 70)
td + geom_point()
td + geom_jitter()
jit <- position_jitter(width = 0.5)
td + geom_jitter(position = jit)
td + geom_jitter(position = jit, colour = alpha("black", 1/10))
td + geom_jitter(position = jit, colour = alpha("black", 1/50))
td + geom_jitter(position = jit, colour = alpha("black", 1/200))
```

## Position adjustments

### Dodge
```{r}
depth_dist <- ggplot(diamonds, aes(clarity))
depth_dist + geom_bar(aes(fill = cut), position = "dodge")
```
Interpretation: We use Dodge to adjust position side by side to compare the values

### Stack
```{r}
depth_dist <- ggplot(diamonds, aes(clarity))
depth_dist + geom_bar(aes(fill = cut), position = "stack")
```
Interpretation: We use Stack to  overlap objects on top of one another to see the total height and see which clarity has the highest count in above graph

### Fill
```{r}
depth_dist <- ggplot(diamonds, aes(clarity)) 
depth_dist + geom_bar(aes(fill = cut), position = "fill")
```
Interpretation: We use Fill to stack for overlapping objects on top of one another and use for comparison of the distribution of cuts within each clarity level in above graph 

### Identity

```{r}
depth_dist <- ggplot(diamonds, aes(clarity))
depth_dist + geom_bar(aes(fill = cut), position = "identity")
```
Interpretation: We use Identity will be plotted directly on top of each other for the same x-axis value. In above case all the cut has been plotted in same x-axis but in s12 clarity the the number of premium is more than ideal so it has come up than the ideal bar.

Task: Ploting the same above code but changing it to histogram and its binwidth to 0.1
```{r}
depth_dist <- ggplot(diamonds, aes(depth)) + xlim(58, 68)
depth_dist + geom_histogram(aes(fill = cut), binwidth = 0.1,
 position = "identity")
```

Task: Using the histogram to plot the depth in x and count in y with cut in stack and using binwidth of 0.1
```{r}
depth_dist <- ggplot(diamonds, aes(depth)) + xlim(58, 68)
depth_dist + geom_histogram(aes(fill = cut), binwidth = 0.1,
 position = "stack")
```

Task: Displaying the histogram with position fill and bindwith to 0.1
```{r}
depth_dist <- ggplot(diamonds, aes(depth)) + xlim(58, 68)
depth_dist + geom_histogram(aes(fill = cut), binwidth = 0.1,
position = "fill")
```

## Overlapping labels

Task: Displaing overlapping labels in categorical varibale
```{r}
ggplot(mpg, aes(x = trans)) +
 geom_bar() +
 labs(x = "Transmission",
 y = "Frequency")

```

Task: Fliping the above bar graph to deal with overlapping issue.
```{r}
# horizontal bar chart
ggplot(mpg, aes(x = trans)) +
 geom_bar() +
 labs(x = "Transmission",
 y = "Frequency") +
 coord_flip()
```

## More Basics plots

Task: Creating boxplot graph
```{r}
ggplot(diamonds,aes(x = cut, y = depth)) +
  geom_boxplot() 
```

Task: Creating jitter graph
```{r}
ggplot(diamonds,
 aes(x = cut,
 y = depth)) +
 geom_jitter()
```

Task: Creating a bubble plot
```{r}
data(mtcars)
ggplot(mtcars,
aes(x = wt, y = mpg, size = hp)) +
geom_point()
```

Task: Creating superheat graph
```{r}
library(superheat)
superheat(mtcars, scale = TRUE)
```

Task: Creating sorted heat map
```{r}
superheat(mtcars,
 scale = TRUE,
 left.label.text.size=3,
 bottom.label.text.size=3,
 bottom.label.size = .05,
 row.dendrogram = TRUE )
```

Task:Creating Basic 3-D scatterplot
```{r}
library(scatterplot3d)
with(mtcars, { scatterplot3d(x = disp,
y = wt,
z = mpg,
main="3-D Scatterplot Example 1")
})
```

Task: Creating 3-D scatterplot graph with blie circles and a line to the horixontal plane
```{r}
with(mtcars, {
scatterplot3d(x = disp,
y = wt,
z = mpg,
color="blue", pch=19,
type = "h",
main = "3-D Scatterplot Example 2",
xlab = "Displacement (cu. in.)",
ylab = "Weight (lb/1000)",
zlab = "Miles/(US) Gallon")
})
```

```{r}
d <- ggplot(diamonds, aes(carat)) + xlim(0, 3)
d + stat_bin(aes(ymax = ..count..), binwidth = 0.1, geom = "area")
```

# Part-II

\begin{enumerate}
\item First thing
\item Second thing
\begin{itemize}
\item A sub-thing
\item Another sub-thing
\end{itemize}
\item Third thing
\end{enumerate}


\begin{itemize}
\item[-] First thing
\item[+] Second thing
\begin{itemize}
\item[Fish] A sub-thing
\item[Plants] Another sub-thing
\end{itemize}
\item[Q] Third thing
\end{itemize}


\documentclass{article}
\title{My First Report}
\author{Add your name in}
\begin{document}
\maketitle
<<echo=FALSE>>=
library (ggplot2)
qplot(carat, data = diamonds, geom = "density", fill = color)
@
hello, world!
\end{document}


