---
title: "Data visualization"
author: "Bibek Sapkota"
output: pdf_document
---

# Data Visualization Lab ggplot2(qplot)

Task 1: Importing ggplot2 library
```{r}
library(ggplot2)
```

## Diamonds Dataset with qplot function

Task: loading the diamonds dataset
```{r}
data("diamonds")
head(diamonds)
```

Task: Using qplot() to produce a scatterplot showing the relationship between the price and carats of a diamond
```{r}
qplot(carat,price,data=diamonds)
```
## Variables Transformation

Task: using log function to transform the variables in a way to make the relationship between variablesless sensitive to outliers.
```{r}
qplot(log(carat), log(price),data=diamonds)
```

## Aesthetic attributes (color,shape,size)

Task: creating a smaller dataset from diaminds by randomly sample. 
```{r}
set.seed(1000)
dsmall<-diamonds[sample(nrow(diamonds),100),]
```

## Color aesthetic

Task: Creating a scatterplot usiing qplot function to show the relationship between the ‘log(carat)’ and ‘log(price)’ of different colours(i.e. ‘color’ variable). 
```{r}
qplot(log(carat),log(price),data=dsmall,colour=color)
```
Task: create a scatter plot to visualize the relationship between ‘log(carat)’ and ‘log(price)’ of different cuts (i.e. ‘cut’
variable).This is by mapping ‘cut’ variable to colour aesthetic attribute.
```{r}
qplot(log(carat),log(price),data = dsmall,colour=cut)
```

## Shape aesthetic

Task:create a scatter plot to visualize the relationship between ‘log(carat)’ and ‘log(price)’ of different cuts(i.e.‘cut’ variable) 
```{r}
qplot(log(carat),log(price),data = dsmall,shape=cut)
```

## Exercise 1

### a. Apply what you have done previously to dsmall dataset with other aesthetic attribute such as‘size’.

```{r}
qplot(log(carat),log(price),data = dsmall,size=clarity)
```

### b. Investigate the relationship between other variables (using colour, shape and size aesthetics) to see if you can find interesting information.

```{r}
qplot(log(depth),log(price),data = dsmall,colour=depth)
```
```{r}
qplot(log(x),log(price),data = dsmall,size=x, colour=x)
```

## Geom 2D

Task: Visualizing the relationship between the natural log of carat and price using a line plot.
```{r}
qplot(log(carat),log(price), data = dsmall,geom="line")
```

## Boxplots ans Jitttered

Task: Visualizing how the price per carat (price/carat) varies across different color grades (color) of diamonds using a jitter plot.
```{r}
qplot(color,price/carat, data = diamonds, geom = "jitter")
```
Task: Visualizing how the price per carat (price/carat) varies across different color grades (color) of diamonds using a boxplot
```{r}
qplot(color,price/carat,data = diamonds,geom = "boxplot")
```

## Geom 1D

Task: Visualizing the distribution of diamond carat using a histogram
```{r}
qplot(carat,data = diamonds,geom = "histogram")
```
Task: Visualizing the distribution of diamond carat  using a density plot
```{r}
qplot(carat,data = diamonds,geom = "density")
```

Task: Visualizing how the distribution of diamond carat varies across different color using a density plot
```{r}
qplot(carat,data = diamonds,geom = "density",colour=color)
```

Task: Visualizing the distribution of diamond carat varies across different color using a filled histogram
```{r}
qplot(carat,data = diamonds,geom = "histogram",fill=color)
```
Task: Visualizing the relationship between diamond carat and price using a scatter plot,
```{r}
qplot(carat,price,data = dsmall,xlab = "Price $", ylab = "Weight (carat)", main = "Price-Weight Relationship")
```

## Exercise 2

### a.Apply bar geoms to show the colour distribution of diamonds in dsmall dataset to see if you can findinteresting relationships.
```{r}
qplot(color,data = dsmall,geom = "bar",xlab="Diamond Color Grade",ylab = "Count",main = "Color Distribution of Diamonds in dsmall")
```

