---
title: "Data visualization Lab"
author: "Bibek Sapkota"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Data Visualization Lab smooting(qplot)

Task 1: Importing ggplot2 library
```{r}
library (ggplot2)
library(splines)
```


Task 2: Importing diamonds dataset and printing it.
```{r}
data <- diamonds
data
```

Task :Creating a random seed to 1000 and create a subset dsmall containing 100 randomly sampled rows from the diamonds dataset.
```{r}
set.seed(1000)
dsmall <- diamonds[sample(nrow(diamonds),100),]
```



Task 3:visualize the relationship between the logarithmically transformed log(carat) and log(price) of diamonds in the dsmall dataset using a scatter plot with a smoothed trend line
```{r}
qplot(log(carat),log(price),data=dsmall,geom = "smooth")
```
Task 4: ploting the logarithm of diamond carat vs. price, displaying both points and a smoothed trend line.
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("point","smooth"))
```
Task 5: Creating a scatter plot of the logarithm of diamond carat versus price from the dsmall dataset, displaying both points and a smoothed trend line with a span of 0.2 for the smoothing parameter. 
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("smooth","point"),span=0.2)
```
Task 6: Creating a scatter plot of the logarithm of diamond carat vs price from the dsmall dataset, displaying both points and a smoothed trend line with a smoothing span parameter set to 1.
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("smooth","point"),span=1)
```

## Fit a linear model instead by using method = “lm” or robust linear model “rlm” which is less sensitive to outlier

Task 1: Visualizing the relationship between the (log(carat)) and (log(price)) of diamonds in the dsmall dataset using a scatter plot with individual points and a linear regression smooth line
```{r}
qplot(log(carat), log(price), data = dsmall, geom = c("point", "smooth"), method = "lm")
```
Task 2: Creating a scatter plot to visualize the relationship between engine displacement (displ) and highway miles per gallon (hwy) for cars in the mpg dataset.
```{r}
qplot(displ,hwy,data = mpg)
```

Task 3: Creating a scatter plot of engine displacement (displ) vs highway miles per gallon (hwy) from the mpg dataset, with points colored by the drive type (drv).
```{r}
qplot(displ, hwy, data = mpg, colour = drv)
```

Task 4: Creating a scatter plot of engine displacement (displ) vs highway miles per gallon (hwy) from the mpg dataset, displaying both points and a smoothed trend line.
```{r}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```
### Question:How are engine size and fuel economy related?

Task 1: creating a scatter plot of engine displacement (displ) versus highway miles per gallon (hwy) from the mpg dataset, with points colored according to the number of cylinders (cyl).
```{r}
qplot(displ,hwy,data = mpg,colour=factor(cyl))
```
task 2: creating a scatter plot of engine displacement (displ) versus highway miles per gallon (hwy) from the mpg dataset, with points colored by the number of cylinders (cyl) and including a linear regression smooth line.
```{r}
qplot(displ,hwy,data = mpg,colour=factor(cyl),geom = c("smooth","point"),method="lm")
```

